---
title: "Exam_ASML-Ex_C-Data1"
output: pdf_document
author: Mohamed ABID
---

```{r "clear all"}
rm(list=ls())   # remove all variables
graphics.off()  # clear all plots
cat('\f')       # clear console
```

# DATA 1

## Loading of libraries

```{r}
library(ggplot2)
library(ggpubr)
library(Matrix)
library(glmnet)
```

## Importation of the data

```{r}
set.seed(3333)
D1 = read.csv('data1.csv', header=TRUE, sep=';')
dim(D1)
```

```{r}
head(D1)
```

```{r}
str(D1)
```

### Convert the type of the qualitative variables

```{r}
D1$Psychologist = as.factor(D1$Psychologist)
D1$Psychologist
```

```{r}
D1$Treatment = as.factor(D1$Treatment)
D1$Treatment
```

### Elimination of the column “Children”

```{r}
D1$Children = NULL
```

## Visualization of the data

```{r}
box.plot.Treatment = ggboxplot(D1, x="Treatment", y="Development"
                               , bxp.errorbar = TRUE
                               , palette = c('red', 'blue') , color = "Treatment")
box.plot.Treatment + theme_bw() + theme(legend.position='None') + rremove("x.grid")
```

```{r}
box.plot.Psychologist = ggboxplot(D1, x='Psychologist', y='Development' 
                                  , bxp.errorbar = TRUE
                                  , palette = c('darkred', 'darkblue') 
                                  , color = 'Psychologist')
box.plot.Psychologist + theme_bw() + theme(legend.position='None') + rremove("x.grid")
```

## Assumption

```{r}
L1.c = lm(Development ~ Treatment + Psychologist + Treatment * Psychologist
          , data=D1)
summary(L1.c)
```

```{r}
plot(L1.c)
```

## ANOVA model

### the complete model:

```{r}
anova(L1.c)
```

### Additive model:

```{r}
L1.a = lm(Development ~ Treatment + Psychologist, data=D1)
anova(L1.a)
```

### Model with only the factor “Treatment”

```{r}
L1.Treatment = lm(Development ~ Treatment, data=D1)
anova(L1.Treatment)
```
```{r}
summary(L1.Treatment)
```
## Step by step strategies: backward strategy

### Backward strategy with AIC criterion 

```{r}
# The linear model with all the explanatory variables and the cross effect
L1.c
```

```{r}
L1.B.AIC = step(L1.c, direction='backward')
```

```{r}
L1.B.AIC
```

### backward strategy with Fisher criterion

```{r}
L1.B.Fisher = step(L1.c, direction='backward', test='F')
```
```{r}
L1.B.Fisher
```

## Step by step strategies: forward strategy

```{r}
L1.1 = lm(D1$Development ~ 1, data=D1)
```

###  forward strategy with AIC criterion

```{r}
L1.F.AIC = step(L1.1, scope=list(L1.1, L1.c), direction='forward')
```
```{r}
L1.F.AIC
```

### forward strategy with Fisher criterion

```{r}
L1.F.Fisher = step(L1.1, scope=list(L1.1, L1.c), direction='forward', test='F')
```
```{r}
L1.F.Fisher
```

## Lasso method

```{r}
X1 = data.matrix(D1[, -3])
Y1 = D1$Development
D1.Lasso = glmnet(X1, Y1, alpha=1)
plot(D1.Lasso, col=c(2, 4), label=TRUE)
```
```{r}
D1.Lasso.cv = cv.glmnet(X1, Y1, alpha=1, nfolds=3)
D1.Lasso.cv
```

```{r}
D1.best.Lambd = D1.Lasso.cv$lambda.1se
D1.best.Lambd
```

```{r}
D1.best.Lasso = glmnet(X1, Y1, alpha=1, lambda=D1.best.Lambd)
coefficients(D1.best.Lasso)
```

## Conclusion

```{r}
L1.best = lm(Development ~ Treatment, data=D1)
L1.best
```
```{r}
summary(L1.best)
```





